Implementing Real-time Transport Services over an Ossified Network

### 2016-05-16 paper.tex
Revise Real-time Transport Services

...discussion of APIs and transport services for real-time applications

- deadlines
- partial reliability
- application-level framing and ADUs
- sub-streams
- etc.

what a conceptual API for this would look like

>>>>

In the IETF, the Transport Services (TAPS) working group is chartered to 
(1) develop a taxonomy of \emph{transport services}. That is, to identify the
features that comprise, and can be combined, to form complete transport
protocols; and (2) to develop an abstract API for applications to request
desirable services, allowing the system to select an appropriate transport
protocol based on application needs. It is hoped that this will loosen the
coupling between application and transport, so enabling deployment of new 
transport protocols.

The work in TAPS is valuable, because it gives us a vocabulary for
discussing the components of transport protocols.  We can then use that
vocabulary when discussing the needs of real-time applications, and the
protocols needed to support those needs.

The essential feature of real-time applications is clearly timing. That is,
since these applications convey real-time data, they all have some concept
of a \emph{deadline} by which that data must be presented, after which it
becomes useless. 
The deadline might be short or long. Interactive applications, such as
telephony, video conferencing, or telepresence, require low end-to-end
latency, with the deadline for presenting the media (playing the audio,
displaying the video frame) being tens, or perhaps low hundreds, of
milliseconds after it was captured. Non-interactive applications, such as
broadcast TV, may have deadlines in the order of seconds, or even tens of
seconds for on-demand programming. 
The deadlines in many networked multimedia applications are unusual when
compared to other real-time systems. They are simultaneously flexible and
strict: flexible in that the exact value of the deadline is typically not
important, provided it is of the right order-of-magnitude for the
application, but strict in that any particular deadline provides a hard
cut-off, after which the data is useless.

The presence of deadlines, in a network that offers only a best-effort
packet deliver service, implies that the transport must provide
\emph{partial reliability}.
The network has non-zero probability of losing any particular packet. 
If forward error correction is used to repair loss, this implies there is
always some probability that a particular packet will be non-recoverable.
If retransmission is used to recover from losses, there is the potential 
for unbounded delay, since any retransmission may itself be lost. 
These probabilities can be estimated, to give probabilistic constraints
on loss and delay, but will be non-zero. Accordingly, to meet deadlines,
the transport must support partial reliability service, accepting that
it may not be able to deliver all data by its deadline.
Many real-time applications run over TCP today, and TCP does not offer a
partial reliability service. This results in occasional play-out stalls, 
when the transport blocks the application because a retransmission takes
too long. These stalls are one of the primary causes of poor user
experience in streaming applications, and are directly caused by the lack
of partial reliability in TCP. Missing one frame that has not been
delivered by its deadline is much less disruptive than a stall in play-out.

Given deadlines and partial reliability, the next important transport
service is \emph{dependency management}. There is no point sending data
that cannot be used, because it depends on previous data that was lost. 
...complex trade-off, since you might want to send data that misses its 
   deadline, if it allows later data to be used.

...following from this, we have the concept of application-level framing
   to make the best use of data, irrespective of losses. This implies that
   the transport provides a \emph{message oriented} service that respects
   ADU boundaries.

...we note that many multimedia applications make use of multiple streams. 
   For example, audio and video are logically separate, etc.
   can be supported by opening multiple transport connections, but
   desirable to support a \emph{sub-stream} service to reduce overheads.

...finally, we note that \emph{congestion control} is important. 
   motivate: historically there has been a belief that real-time doesn't 
   do congestion control, needs an isochronous channel. 
   not practical for the Internet. 
   Also, experience has shown not necessary for many real-time applications.
   Yes, there are some that need constant bandwidth and cannot adapt, but
   many are adaptive. 
   Better user experience to adapt to fit the channel, rather than recover
   from loss after the channel discards packets that don't fit.

..what a conceptual API for this would look like?
